---
title: Key literature on _Schistosoma_ and host-parasite systems    
author: |
 | Matthew Malishev^1^*  
 |  
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: blue
urlcolor: blue
citecolor: red
output:
  html_document:
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 3
    number_sections: no
    toc_float: yes
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 3
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Emory/admin/postdocs/postdoc_meetups"
  date: !r Sys.Date()
  session: !r sessionInfo()  
  version: !r getRversion()
  email: "matthew.malishev@emory.edu"
  doi: https://github.com/darwinanddavis/
classoption: portrait
# ^['https://github.com/darwinanddavis/UsefulCode'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=1000)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=1000), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\newpage   

Date: `r params$date`  
R version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`  

\newpage  

## Overview

Key literature and areas of research in _Schistosoma_ and host-parasite systems.  

## Notes  

**Models with schisto**  

Aim: To account for time lag in population cycles based on things like resources and size structure  

- Charles King  
- Mark EJ Woolhouse (1991, 1992)  
- David Rollinson   

\newpage  

*****  
## Habitat and resources  

[Southgate VR (1997) Schistosomiasis in the Senegal River Basin: Before and after the construction of the dams at Diama, Senegal and Manantali, Mali and future pros- pects. J Helminthol 71(2):125–132.](Southgate VR (1997) Schistosomiasis in the Senegal River Basin: Before and after the construction of the dams at Diama, Senegal and Manantali, Mali and future pros- pects. J Helminthol 71(2):125–132)    

> From Sokolow et al 2015 PNAS   

\newpage  

*****  
## Human health stats  

[Steinmann P, Keiser J, Bos R, Tanner M, Utzinger J (2006) Schistosomiasis and water resources development: Systematic review, meta-analysis, and estimates of people at risk. Lancet Infect Dis 6(7):411–425]()

> From Sokolow et al 2015 PNAS  
> Schistosomiasis infects an estimated 220–240 million people globally, and 790 million are at risk for infection, more than 90% of whom are in Sub-Saharan Africa (14).  

[WHO (2015) Preventive Chemotherapy and Transmission Control (PCT) databank](www.who.int/neglected_diseases/preventive_chemotherapy/databank/en/)  
[World Health Assembly (2012) Elimination of Schistosomiasis in WHA65/2012/REC/1 Sixty-Fifth World Health Assembly: Resolutions and Decisions Annexes (WHO, Geneva)]()  

[Cheever AW, Macedonia JG, Mosimann JE, Cheever EA (1994) Kinetics of egg pro- duction and egg excretion by Schistosoma mansoni and S. japonicum in mice infected with a single pair of worms. Am J Trop Med Hyg 50(3):281–295]()  

> From Sokolow et al 2015 PNAS  
> Each infected snail sheds thousands of cer- cariae, which seek and penetrate human skin. After entering the skin, the parasites migrate to the blood vessels of the intestines (S. mansoni) or urinary bladder (S. hematobium), where female worms lay 350–2,200 eggs per day (15)  

[Jobin WR, Negrón-Aponte H, Michelson EH (1976) Schistosomiasis in the Gorgol Valley of Mauritania. Am J Trop Med Hyg 25(4):587–594]()  

> From Sokolow et al 2015 PNAS  

> Death from liver failure or bladder cancer can be preceded by chronic anemia, cognitive im- pairment in children, growth stunting, infertility, and a higher risk of contracting HIV in women (17, 18).  

[Whole genome sequencing and morphological analysis of the human-infecting schistosome emerging in Europe reveals a complex admixture between Schistosoma haematobium and Schistosoma bovis parasites.](https://www.biorxiv.org/content/early/2018/08/11/387969)    

[Hotez PJ, Fenwick A, Kjetland EF (2009) Africa’s 32 cents solution for HIV/AIDS. PLoS Negl Trop Dis 3(5):e430]()  

> Notes  

[WHO (2015) Preventive Chemotherapy and Transmission Control (PCT) databank.](www.who.int/neglected_diseases/preventive_chemotherapy/databank/)  

> Notes  

[WHO (2011) Schistosomiasis: Progress Report 2001-2011 and Strategic Plan 2012-2020 (WHO, Geneva]().  

> Notes  

[Bockarie MJ, Kelly-Hope LA, Rebollo M, Molyneux DH (2013) Preventive chemother- apy as a strategy for elimination of neglected tropical parasitic diseases: Endgame challenges. Philos Trans R Soc Lond B Biol Sci 368(1623):20120144.]()  

> Notes  

[Gray DJ, et al. (2010) Schistosomiasis elimination: Lessons from the past guide the future. Lancet Infect Dis 10(10):733–736]()  

> Notes  

[Fenwick A, Savioli L (2011) Schistosomiasis elimination. Lancet Infect Dis 11(5):346, author reply 346–347]()  

> Notes  

[Zhang Z, Jiang Q (2011) Schistosomiasis elimination. Lancet Infect Dis 11(5):345, au- thor reply 346–347]()  

> Notes  

[World Health Assembly (2012) Elimination of Schistosomiasis in WHA65/2012/REC/1 Sixty-Fifth World Health Assembly: Resolutions and Decisions Annexes (WHO, Geneva).]()  

> Notes  

\newpage  

*****
## Starvation  

[Nelson et al. 2016 Effects of abnormal temperature and starvation on the internal defense system of the schistosome-transmitting snail Biomphalaria glabrata](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4945359/)  

> Notes  

[Andre´ Gergs and Tjalling Jager 2014 Body size-mediated starvation resistance in an insect predator](https://besjournals.onlinelibrary.wiley.com/doi/pdf/10.1111/1365-2656.12195)  

> Notes  

\newpage  

*****
## Molluscicide    

[Sokolow et al 2018 To Reduce the Global Burden of Human Schistosomiasis, Use ‘Old Fashioned’ Snail Control]()  

> Contains table of papers looking at schisto control programs and strategies     

\newpage  

*****
## Immunity  


[Cressler et al 2014 Disentangling the interaction among host resources, the immune system and pathogens](https://onlinelibrary.wiley.com/doi/pdf/10.1111/ele.12229)  

> Notes  

\newpage  

*****  
## Predation  

[Sokolow et al 2015 Reduced transmission of human schistosomiasis after restoration of a native river prawn that preys on the snail intermediate host](http://www.pnas.org/content/pnas/early/2015/07/15/1502651112.full.pdf)  

> 'Reinfection after treatment is a problem that plagues efforts to control parasites with complex transmission pathways, such as schistosomiasis, which affects at least 220 million people worldwide and requires an obligate snail intermediate host.'     

> 





