---
title: Instructions for runnng the individual-based Dynamic Energy Budget _Schistosoma_ population model (SIDEB) on Windows     
author: |
 | Matthew Malishev^1^* & David J. Civitello^1^   
 |  
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: pink
urlcolor: blue
citecolor: red
output:
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  html_document:
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Emory/research/schisto_ibm/DEB_IBM/docs/DEB_IBM_readme"
  date: !r Sys.Date()
  session: !r sessionInfo()  
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/SchistoIBM/tree/master/windows
classoption: portrait
# ^['https://github.com/darwinanddavis/SchistoIBM/tree/master/windows'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\newpage   

Date: `r params$date`  
R version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`  

\  

R session info 

```{r echo=F}
params$session
```      

\newpage  

## Overview

Follow the instructions to run the simulation model from `R` or `RStudio`. All reports and bugs should be addressed to matthew.malishev@gmail.com.   

## Required files    
Files required for running the simulation are outlined below and will be automatically loaded from the [*Schistosoma* IBM Github page](https://github.com/darwinanddavis/SchistoIBM) when running the model:  

```
 DEB_IBM.R  
 DEB_INF_GUTS_IBM.nlogo  
 FullStarve_shrink_production2.Rda  
 IndividualModel_IBM.c  
 IndividualModel_IBM.dll (Windows, generated from C)      
```
## Install R RStudio
  
Install `RStudio` from the `RStudio` [website](https://www.rstudio.com/products/rstudio/download/#download).  

## Install NetLogo  

Install NetLogo from the NetLogo [website](https://ccl.northwestern.edu/netlogo/download.shtml).  

## Run `RNetLogo`    

Load the 'DEB_IBM.R' file into your `R` session. Follow the instructions to load the model and execute the simulation.  

## References  

^1^ [Installing R Studio](https://www.rstudio.com/products/rstudio/download/#download)   

^2^ [Installing NetLogo](https://www.rstudio.com/products/rstudio/download/#download)    








