---
title: Instructions for runnng the individual-based Dynamic Energy Budget _Schistosoma_ population model (SIDEB) on Mac OSX     
author: |
 | Matthew Malishev^1^* & David J. Civitello^1^   
 |  
 | _^1^ Department of Biology, Emory University, 1510 Clifton Road NE, Atlanta, GA, USA, 30322_
#bibliography:/Users/malishev/Documents/Melbourne Uni/Thesis_2016/library.bib
fontsize: 10
geometry: margin=1in
documentclass: article
linkcolor: blue  
urlcolor: blue
citecolor: red
output:
  pdf_document:
    includes:
      in_header: # add .tex file with header content
    highlight: tango
    template: null
    toc: yes
    toc_depth: 4
    number_sections: false
    fig_width: 4
    fig_height: 5
    fig_caption: true
    df_print: tibble 
    citation_package: biblatex # natbib
    latex_engine: xelatex #pdflatex # lualatex
    keep_tex: true # keep .tex file in dir 
  html_document:
    highlight: tango
    code_folding: hide
    toc: yes
    toc_depth: 4
    number_sections: no
    toc_float: yes
  word_document:
    highlight: tango
    keep_md: yes
    pandoc_args: --smart
    #reference: mystyles.docx
    toc: yes
inludes:
  before_body: before_body.tex
subtitle: 
tags:
- nothing
- nothingness
params: 
  dir: "/Users/malishev/Documents/Emory/research/schisto_ibm/SchistoIBM/mac"
  date: !r Sys.Date()
  session: !r sessionInfo()  
  version: !r getRversion()
  email: "matthew.malishev@gmail.com"
  doi: https://github.com/darwinanddavis/SchistoIBM/tree/master/mac
classoption: portrait
# ^['https://github.com/darwinanddavis/SchistoIBM/tree/master/mac'] # footnote
vignette: >
  %\VignetteIndexEntry{Useful R code}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

```{r echo = FALSE}
# library(rmarkdown)
# setwd("")
# f <- list.files()[1]
# render(f, output_format='pdf_document')
```

```{r, set-options, echo = FALSE, cache = FALSE}
options(width=100)
knitr::opts_chunk$set(
 eval = TRUE, # run all code
 # echo = FALSE, # show code chunks in output 
 comment = "",
 tidy.opts=list(width.cutoff=100), # set width of code chunks in output
 tidy=TRUE, # make output as tidy
 message = FALSE,  # mask all messages
 warning = FALSE, # mask all warnings 
 size="small" # set code chunk size
)

# https://github.com/ucb-stat133/stat133-fall-2016/blob/master/hws/hw02-tables-ggplot.Rmd
knitr::opts_knit$set(root.dir=paste0(params$dir,"/")) # set working dir

setwd(paste0(params$dir,"/")) # for running just in R not knitr
```

\newpage   

Date: `r params$date`  
R version: `r params$version`  
*Corresponding author: `r params$email`  
This document can be found at `r params$doi`  

\  

R session info 

```{r echo=F}
params$session
```      

\newpage  

## Overview  

Follow the instructions to run the simulation model from `R` or `RStudio`. All reports and bugs should be addressed to matthew.malishev@gmail.com.   

## Required files  
Files required for running the simulation are outlined below and will be automatically loaded from the [*Schistosoma* IBM Github page](https://github.com/darwinanddavis/SchistoIBM) when running the model:  

```
DEB_IBM.R
DEB_INF_GUTS_IBM.nlogo
FullStarve_shrink_production2.Rda
IndividualModel_IBM.c
IndividualModel_IBM.so
IndividualModel_IBM.o
```

## Set system to run C toolchain  
1. Follow the steps for compiling a toolchain and running C code in `R` outlined at [Installing compiler toolchain for Mac OSX](https://thecoatlessprofessor.com/programming/r-compiler-tools-for-rcpp-on-macos/). See also [[4]](#references) if the above steps don't work. Thanks to the [www.thecoatlessprofessor.com](www.thecoatlessprofessor.com).         

## Java tests  
Diagnostics for testing you have the correct version of Java on your computer. Tests are sequential (from [[5]](#references)).     

1. Open `R` and run the following code.  

```{r, eval=F, results="hide"}
# test java is working
require(RCurl)
script <- getURL("https://raw.githubusercontent.com/darwinanddavis/SchistoIBM/master/mac/java_test.R", ssl.verifypeer = FALSE)
eval(parse(text = script))
capture.output(errorlist()) # read error message from java_test.R output

```

If you get an `R` error message e.g. `Error in .jnew ...` or an error beginning with `Failed Test ...`, move onto ["Install JGR"](#install-jgr). Otherwise, continue to [Run RNetLogo](#run-rnetlogo).    

## Run `RNetLogo`  

Load the 'DEB_IBM.R' file into your `R` session. Follow the instructions to load the model and execute the simulation.  

If you see the following `rJava` error in your `R` session, proceed to step [Install `rJava`](#install-rjava):  
\  
\  
```
Error : .onLoad failed in loadNamespace() for 'rJava', details:
  call: dyn.load(file, DLLpath = DLLpath, ...)
  error: unable to load shared object ...
  .  
  .  
  .  
```
\  
\  

## Install `rJava`  

Open Terminal ($Applications > Terminal$) and run the following in Terminal (sourced from [[1]](#references) and [[2]](#references)) 
\  
```{bash, eval=F}
sudo ln -s $(/usr/libexec/java_home)/jre/lib/server/libjvm.dylib /usr/local/lib
```
\  
Now re-run the 'DEB_IBM.R' code in `R`.  

The following error in `R` when executing `NLStart()` means you have successfully navigated the `rJava` errors, but are running into a Java issue that's preventing `R` from generating the NetLogo GUI from `R`:
\  
\  
```
java.awt.HeadlessException
    at java.awt.GraphicsEnvironment.checkHeadless(GraphicsEnvironment.java:204)
    at java.awt.Window.<init>(Window.java:536)
    at java.awt.Frame.<init>(Frame.java:420)
    at java.awt.Frame.<init>(Frame.java:385)
    at javax.swing.SwingUtilities$SharedOwnerFrame.<init>(SwingUtilities.java:1758)  
 .   
 .  
 .  
```  
\  
For the above error, the `NLStart()` function should run successfully with `gui=F`, which launches a headless GUI mode.
\  
```{r eval=F}
NLStart(nl.path,gui=F,nl.jarname = paste0("netlogo-",ver_nl,".jar")) # open netlogo w/o gui
```

If the above steps don't work, move onto [Install JGR](#install-jgr).      

## Install JGR  
\  
1. Check you have the latest version of [Java/Oracle](https://www.oracle.com/technetwork/java/javase/downloads/index-jsp-138363.html).  

2. Install JGR by running the following code in your `R` session (from [[3]](#references)):
\  
```{r eval=F, results="hide"}
install.packages('JGR',,'http://www.rforge.net/')
library(JGR)
JGR::JGR()
```

Now load the 'DEB_IBM.R' file into your `JGR` session to run the simulation model.   

For up to date troubleshooting running `RNetLogo` on `R` v. 6.0 or higher, see [Running Netlogo 6.0.+](https://github.com/NetLogo/NetLogo/issues/1282).  

## References  
^1^ [rJava load error in RStudio/R after “upgrading” to OSX Yosemite](https://stackoverflow.com/questions/30738974/rjava-load-error-in-rstudio-r-after-upgrading-to-osx-yosemite)      

^2^ [Run rJava with RStudio using OSX 10.10](http://paulklemm.com/blog/2015-02-20-run-rjava-with-rstudio-under-osx-10-dot-10/)  

^3^ [Installing JGR](https://github.com/markush81/JGR) 

^4^ [GCC compiler in R](https://stackoverflow.com/questions/1616983/building-r-packages-using-alternate-gcc)  

^5^ [Test for working version of Java on your computer](https://stackoverflow.com/questions/14915898/rnetlogo-function-nlstart-fails-to-launch-gui)  
